The Impact of Past Behavior Normality on Regret: A Replication Study

### 2024-02-15 outputs/paper/Paper_2.qmd
update

To further explore this concept, we first check for figure1, it's apparent that in the survey, more people believe Mr. Jones, who rarely offers rides, would feel a stronger sense of regret for being robbed due to offering a ride just once. Firstly, the habit of offering rides initially influences one's psychological expectations. For instance, someone who has never offered rides before would tend to believe that their kind act, or an exception, would to some extent help others or receive positive psychological feedback. However, conversely, when their psychological expectations are not met, the sense of psychological disparity comes into play, leading people to feel that acts of kindness might instead invite unnecessary trouble, thereby intensifying their psychological fixation. Positive feedback would make them happier, while negative impacts would bring about more feelings of regret.

From a psychological perspective，counterfactual thinking involves individuals imagining how different actions could have led to different outcomes, particularly after experiencing adverse results. This mental process becomes especially pronounced when attempting new experiences and encountering disasters, as people tend to evaluate their decisions by constructing scenarios that differ from what actually happened. For instance, consider someone who decides to engage in a high-risk investment, hoping for significant returns, but instead faces substantial financial losses. In this context, counterfactual thinking might manifest as thoughts like, "If I hadn't invested in that project, my financial situation would be better now." Such thinking not only deepens the individual's sense of regret but might also lead to doubts about their decision-making abilities. Given that individuals tend to assess the positive outcomes of unchosen alternatives, this mode of thinking amplifies regret over the actual choice made. Therefore, counterfactual thinking reveals how people enhance their feelings of regret by comparing actual outcomes with imagined, potentially better ones, especially when venturing into new experiences that yield unexpected negative consequences.

we secondly need to verify whether accidents in exceptional behaviors lead to more regret compared to accidents in routine behaviors. Venturing into new experiences and facing setbacks often leads to heightened feelings of regret, a phenomenon deeply rooted in our psychological responses to deviations from routine. When individuals engage in behaviors that stray from their norm—deciding on a whim to take a different route home, trying out an unconventional restaurant, or making an impulsive purchase—any negative outcomes tend to be felt more acutely. This intensified regret is linked to the concept of "exceptionality," where actions that stand out as exceptions to regular patterns are more likely to trigger stronger counterfactual thinking. Essentially, when outcomes don't go as planned, our minds more easily imagine alternative scenarios where sticking to the routine could have avoided the undesired result.

@@

From the article, we can see that among 277 people, 81% believe that a person named White who was involved in a similar accident while driving on a route he only took for a change of scenery would feel greater regret and worse luck. Therefore, the factor of luck often has a direct relationship with the level of regret. When someone encounters a random event, the worse their luck, the stronger their feeling of regret. For example, Mr. Adams had an accident while driving home on his usual route after work. Mr. White had a similar accident while he was driving on a route he only took when he wanted a change of scenery. In such cases, for accidents that happen under normal circumstances, people might think that it was something bound to happen, believing that no matter what, if they were on that road, they would have faced the same situation, and thus, they wouldn't particularly feel unlucky or have much regret. However, people tend to attribute accidents that happen under unusual circumstances to bad luck. Mr. White's temporary choice led to a bad outcome. Since there was no way to weigh the probabilities of different choices at the moment of decision, people are more likely to attribute the event to bad luck, resulting in a stronger sense of regret.

The way people understand luck in relation to regret is detailed and subtle. Essentially, luck is about the element of chance in what happens after we make decisions or take actions. When people think that what happens is down to luck, their feelings of regret can change depending on how much control they think they had over the situation. If someone believes a bad outcome was just unlucky, they might not feel as much regret because they see it as something they couldn't avoid. On the other hand, if they see the outcome as the result of their own choices, the regret can be stronger because they think they could have chosen differently to prevent the bad result. This shows that how much regret people feel isn't just about what happens, but also how they interpret these happenings in terms of luck and their own actions.

@@

The interplay between counterfactual thinking and the consequent feelings of regret significantly influences individuals' expectations for compensation, a dynamic vividly illustrated through the experiences of White and Smith. Counterfactual thinking, the cognitive process of envisioning alternative scenarios that did not occur, can intensify feelings of regret, especially when these imagined outcomes are perceived as more favorable than the actual events. For White, who frequented a familiar store and experienced a robbery resulting in severe injury, the compensation expectations are densely clustered around $500,000, suggesting that the routine nature of his visit might have led to less intense counterfactual thinking and subsequently, a more contained range of expected compensation. In contrast, Smith, who ventured into an unfamiliar store under unusual circumstances and encountered the same fate, experiences stronger regret, fueled by more potent counterfactual thoughts such as "If only I had avoided making that unusual choice." This heightened level of regret is reflected in the more dispersed compensation expectations, with most believing that the figure should lie between ¥500,000 and ¥1,000,000. This scenario underlines how the depth and focus of counterfactual thinking can dramatically shape an individual's perception of their loss and their valuation of necessary compensation. It showcases the complex relationship between cognitive processes and emotional responses, demonstrating that individuals' compensation expectations are not only a reflection of their physical losses but also an attempt to mitigate the emotional distress caused by their reflections on what could have been, thus highlighting the profound impact of counterfactual thinking on human behavior and expectations in the aftermath of traumatic events.

>>>>

To further explore this concept, we first check for figure1, it's apparent that in the survey, more people believe Mr. Jones, who rarely offers rides, would feel a stronger sense of regret for being robbed due to offering a ride just once. Firstly, the habit of offering rides initially influences one's psychological expectations. For instance, someone who has never offered rides before would tend to believe that their kind act, or an exception, would to some extent help others or receive positive psychological feedback. However, conversely, when their psychological expectations are not met, the sense of psychological disparity comes into play, leading people to feel that acts of kindness might instead invite unnecessary trouble, thereby intensifying their psychological fixation. Positive feedback would make them happier, while negative impacts would bring about more feelings of regret (Kutscher & Feldman 2019).

From a psychological perspective，counterfactual thinking involves individuals imagining how different actions could have led to different outcomes, particularly after experiencing adverse results. This mental process becomes especially pronounced when attempting new experiences and encountering disasters, as people tend to evaluate their decisions by constructing scenarios that differ from what actually happened (Roese & Olson, 1997). For instance, consider someone who decides to engage in a high-risk investment, hoping for significant returns, but instead faces substantial financial losses. In this context, counterfactual thinking might manifest as thoughts like, "If I hadn't invested in that project, my financial situation would be better now." Such thinking not only deepens the individual's sense of regret but might also lead to doubts about their decision-making abilities. Given that individuals tend to assess the positive outcomes of unchosen alternatives, this mode of thinking amplifies regret over the actual choice made. Therefore, counterfactual thinking reveals how people enhance their feelings of regret by comparing actual outcomes with imagined, potentially better ones, especially when venturing into new experiences that yield unexpected negative consequences (Roese & Olson, 1997).

we secondly need to verify whether accidents in exceptional behaviors lead to more regret compared to accidents in routine behaviors. Venturing into new experiences and facing setbacks often leads to heightened feelings of regret, a phenomenon deeply rooted in our psychological responses to deviations from routine. When individuals engage in behaviors that stray from their norm—deciding on a whim to take a different route home, trying out an unconventional restaurant, or making an impulsive purchase—any negative outcomes tend to be felt more acutely. This intensified regret is linked to the concept of "exceptionality," where actions that stand out as exceptions to regular patterns are more likely to trigger stronger counterfactual thinking (Briazu et al, 2017). Essentially, when outcomes don't go as planned, our minds more easily imagine alternative scenarios where sticking to the routine could have avoided the undesired result.

@@

From the article, we can see that among 277 people, 81% believe that a person named White who was involved in a similar accident while driving on a route he only took for a change of scenery would feel greater regret and worse luck. Therefore, the factor of luck often has a direct relationship with the level of regret. When someone encounters a random event, the worse their luck, the stronger their feeling of regret. For example, Mr. Adams had an accident while driving home on his usual route after work. Mr. White had a similar accident while he was driving on a route he only took when he wanted a change of scenery. In such cases, for accidents that happen under normal circumstances, people might think that it was something bound to happen, believing that no matter what, if they were on that road, they would have faced the same situation, and thus, they wouldn't particularly feel unlucky or have much regret (Kutscher & Feldman 2019). However, people tend to attribute accidents that happen under unusual circumstances to bad luck. Mr. White's temporary choice led to a bad outcome. Since there was no way to weigh the probabilities of different choices at the moment of decision, people are more likely to attribute the event to bad luck, resulting in a stronger sense of regret.

The way people understand luck in relation to regret is detailed and subtle. Essentially, luck is about the element of chance in what happens after we make decisions or take actions. When people think that what happens is down to luck, their feelings of regret can change depending on how much control they think they had over the situation. If someone believes a bad outcome was just unlucky, they might not feel as much regret because they see it as something they couldn't avoid. On the other hand, if they see the outcome as the result of their own choices, the regret can be stronger because they think they could have chosen differently to prevent the bad result (Kutscher & Feldman 2019). This shows that how much regret people feel isn't just about what happens, but also how they interpret these happenings in terms of luck and their own actions.

@@

The interplay between counterfactual thinking and the consequent feelings of regret significantly influences individuals' expectations for compensation, a dynamic vividly illustrated through the experiences of White and Smith. Counterfactual thinking, the cognitive process of envisioning alternative scenarios that did not occur, can intensify feelings of regret, especially when these imagined outcomes are perceived as more favorable than the actual events. For White, who frequented a familiar store and experienced a robbery resulting in severe injury, the compensation expectations are densely clustered around $500,000, suggesting that the routine nature of his visit might have led to less intense counterfactual thinking and subsequently, a more contained range of expected compensation. In contrast, Smith, who ventured into an unfamiliar store under unusual circumstances and encountered the same fate, experiences stronger regret, fueled by more potent counterfactual thoughts such as "If only I had avoided making that unusual choice." This heightened level of regret is reflected in the more dispersed compensation expectations, with most believing that the figure should lie between ¥500,000 and ¥1,000,000. This scenario underlines how the depth and focus of counterfactual thinking can dramatically shape an individual's perception of their loss and their valuation of necessary compensation. It showcases the complex relationship between cognitive processes and emotional responses, demonstrating that individuals' compensation expectations are not only a reflection of their physical losses but also an attempt to mitigate the emotional distress caused by their reflections on what could have been, thus highlighting the profound impact of counterfactual thinking on human behavior and expectations in the aftermath of traumatic events (Kutscher & Feldman 2019).
